```{r, include=F, echo=F}

source("../init_settings.R")

library(rethinking)
data(chimpanzees)
d <- chimpanzees

rm(chimpanzees)
detach(package:rethinking, unload = T)
library(brms)

```

Dataset information
* actor : name of actor
* recipient : name of recipient (NA for partner absent condition)
* condition : partner absent (0), partner present (1)
* block : block of trials (each actor x each recipient 1 time)
* trial : trial number (by chimp = ordinal sequence of trials for each chimp, ranges from 1-72; partner present trials were interspersed with partner absent trials)
* prosocial_left : 1 if prosocial (1/1) option was on left; 0 === 2 food items present on right side & 1 === prosocail option present on left
* chose_prosoc : choice chimp made (0 = 1/0 option, 1 = 1/1 option)
* pulled_left : which side did chimp pull (1 = left, 0 = right)

We are trying to study if the chimpanzee pulled the left level when food as present and a pro-social option existed.

There are four combinations:
(1) prosoc_left= 0 and condition= 0: Two food items on right and no partner.
(2) prosoc_left= 1 and condition= 0: Two food items on left and no partner.
(3) prosoc_left= 0 and condition= 1: Two food items on right and partner present.
(4) prosoc_left= 1 and condition= 1: Two food items on left and partner presen

```{r}

d$treatment <- 1 + d$prosoc_left + 2 * d$condition
d
```

The question that we will now try to model here is whether the left level was pulled based on the treatment and the actor.

L ~ Binomial(1, p)
p ~ actor[i] = treatment[i]

Setting the priors for the model
```{r}

get_prior(data = d, formula = pulled_left | trials(1) ~ 1, family =binomial)

```

An intercept only simple binomial model

```{r}

mfile <- "models/chimp_bm_test1"

# file.remove(paste0(mfile, ".rds"))

priors <- prior(normal(0,10), class = Intercept)

bm_test1 <- brm(data = d, family = binomial,
          formula = pulled_left | trials(1) ~ 1,
          sample_prior = T, prior = priors,
          cores = 4, chains = 4,
          file = mfile)

mfile <- "models/chimp_bm_test2"

# file.remove(paste0(mfile, ".rds"))

priors <- prior(normal(0,1.5), class = Intercept)

bm_test2 <- brm(data = d, family = binomial,
          formula = pulled_left | trials(1) ~ 1,
          sample_prior = T, prior = priors,
          cores = 4, chains = 4,
          file = mfile)

```

Priot to building the model, what does the model think is plausible?
```{r}

p1 <- prior_samples(bm_test1) %>% 
  inv_logit_scaled() %>%
  ggplot(aes(x = Intercept)) + geom_density()

p2 <- prior_samples(bm_test2) %>% 
  inv_logit_scaled() %>%
  ggplot(aes(x = Intercept)) + geom_density()

grid.arrange(p1, p2, nrow = 1)
```

Adding the treatment column

Identifying the priors
```{r}

get_prior(family = binomial, data = d, formula = pulled_left | trials(1) ~ 1 + treatment)

```


```{r}

mfile <- "models/chimp_m11.2"

# file.remove(paste0(mfile, ".rds"))

priors <- c(prior(normal(0, 1.5), class = Intercept),
            prior(normal(0, 10), class = b, coef = "treatment"))

bm_11.2 <- brm(data = d, family = binomial,
          formula = pulled_left | trials(1) ~ 1 + treatment,
          sample_prior = T, prior = priors,
          cores = 4, chains = 4,
          file = mfile)


mfile <- "models/chimp_m11.3"

# file.remove(paste0(mfile, ".rds"))

priors <- c(prior(normal(0, 1.5), class = Intercept),
            prior(normal(0, 0.5), class = b, coef = "treatment"))

bm_11.3 <- brm(data = d, family = binomial,
          formula = pulled_left | trials(1) ~ 1 + treatment,
          sample_prior = T, prior = priors,
          cores = 4, chains = 4,
          file = mfile)

```

```{r}

p1 <- prior_samples(bm_11.2) %>%
  mutate(preds = inv_logit_scaled(b + Intercept)) %>%
  ggplot(aes(x = preds)) + geom_density()
  
p2 <- prior_samples(bm_11.3) %>%
  mutate(preds = inv_logit_scaled(b + Intercept)) %>%
  ggplot(aes(x = preds)) + geom_density()

grid.arrange(p1, p2, nrow =1)

```

Considering both actor and treatment

```{r}

mfile <- "models/chimp_m11.4"

# file.remove(paste0(mfile, ".rds"))

priors <- c(prior(normal(0, 1.5), class = Intercept),
            prior(normal(0, 0.5), class = b, coef = "treatment"),
            prior(normal(0, 1.5), class = b, coef = "actor"))

bm_11.4 <- brm(data = d, family = binomial,
          formula = pulled_left | trials(1) ~ 1 + actor + treatment,
          sample_prior = T, prior = priors,
          cores = 4, chains = 4,
          file = mfile)

```

```{r}


```