```{r, include=F, echo=F}

source("../init_settings.R")

library(rethinking)
data(chimpanzees)
d <- chimpanzees

rm(chimpanzees)
detach(package:rethinking, unload = T)
library(brms)

```

Dataset information
* actor : name of actor
* recipient : name of recipient (NA for partner absent condition)
* condition : partner absent (0), partner present (1)
* block : block of trials (each actor x each recipient 1 time)
* trial : trial number (by chimp = ordinal sequence of trials for each chimp, ranges from 1-72; partner present trials were interspersed with partner absent trials)
* prosocial_left : 1 if prosocial (1/1) option was on left; 0 === 2 food items present on right side & 1 === prosocail option present on left
* chose_prosoc : choice chimp made (0 = 1/0 option, 1 = 1/1 option)
* pulled_left : which side did chimp pull (1 = left, 0 = right)

We are trying to study if the chimpanzee pulled the left level when food as present and a pro-social option existed.

There are four combinations:
(1) prosoc_left= 0 and condition= 0: Two food items on right and no partner.
(2) prosoc_left= 1 and condition= 0: Two food items on left and no partner.
(3) prosoc_left= 0 and condition= 1: Two food items on right and partner present.
(4) prosoc_left= 1 and condition= 1: Two food items on left and partner presen

```{r}

d$treatment <- 1 + d$prosoc_left + 2 * d$condition
d$treatment <- as.factor(d$treatment)
d$pulled_left <- as.factor(d$pulled_left)
d$actor <- as.factor(d$actor)
d$block <- as.factor(d$block)

d
```

The question that we will now try to model here is whether the left level was pulled based on the treatment and the actor.

L ~ Binomial(1, p)
p ~ actor[i] = treatment[i]

Setting the priors for the model
```{r}

get_prior(data = d, formula = pulled_left | trials(1) ~ 1, family =binomial)

```

An intercept only simple binomial model

```{r}

mfile <- "models/chimp_bm_test1"

# file.remove(paste0(mfile, ".rds"))

priors <- prior(normal(0,10), class = Intercept)

bm_test1 <- brm(data = d, family = binomial,
          formula = pulled_left | trials(1) ~ 1,
          sample_prior = T, prior = priors,
          cores = 4, chains = 4,
          file = mfile)

mfile <- "models/chimp_bm_test2"

# file.remove(paste0(mfile, ".rds"))

priors <- prior(normal(0,1.5), class = Intercept)

bm_test2 <- brm(data = d, family = binomial,
          formula = pulled_left | trials(1) ~ 1,
          sample_prior = T, prior = priors,
          cores = 4, chains = 4,
          file = mfile)

```

Priot to building the model, what does the model think is plausible?
```{r}

p1 <- prior_samples(bm_test1) %>% 
  inv_logit_scaled() %>%
  ggplot(aes(x = Intercept)) + geom_density()

p2 <- prior_samples(bm_test2) %>% 
  inv_logit_scaled() %>%
  ggplot(aes(x = Intercept)) + geom_density()

grid.arrange(p1, p2, nrow = 1)
```

Adding the treatment column

Identifying the priors
```{r}

get_prior(family = binomial, data = d, formula = pulled_left | trials(1) ~ 1 + treatment)

```


```{r}

mfile <- "models/chimp_m11.2"

# file.remove(paste0(mfile, ".rds"))

get_prior(data = d, family = binomial,
          formula = pulled_left | trials(1) ~ 1 + treatment)

priors <- c(prior(normal(0, 1.5), class = Intercept),
            prior(normal(0, 10), class = b))

bm_11.2 <- brm(data = d, family = bernoulli,
          formula = pulled_left ~ 1 + treatment,
          sample_prior = T, prior = priors,
          cores = 4, chains = 4,
          file = mfile)


mfile <- "models/chimp_m11.3"
# file.remove(paste0(mfile, ".rds"))

priors <- c(prior(normal(0, 1.5), class = Intercept),
            prior(normal(0, 0.5), class = b))

bm_11.3 <- brm(data = d, family = bernoulli,
          formula = pulled_left ~ 1 + treatment,
          sample_prior = T, prior = priors,
          cores = 4, chains = 4,
          file = mfile)

```

```{r}

p1 <- prior_samples(bm_11.2) %>%
  mutate(preds = inv_logit_scaled(b + Intercept)) %>%
  ggplot(aes(x = preds)) + geom_density()
  
p2 <- prior_samples(bm_11.3) %>%
  mutate(preds = inv_logit_scaled(b + Intercept)) %>%
  ggplot(aes(x = preds)) + geom_density()

grid.arrange(p1, p2, nrow =1)

```

Considering both actor and treatment

```{r}


# priors <- c(prior(normal(0, 0.5), class = b, coef = treatment2),
#              prior(normal(0, 0.5), class = b, coef = treatment3),
#              prior(normal(0, 0.5), class = b, coef = treatment4),
#              prior(normal(0, 1.5), class = b, coef = actor1),
#             prior(normal(0, 1.5), class = b, coef = actor2),
#             prior(normal(0, 1.5), class = b, coef = actor3),
#             prior(normal(0, 1.5), class = b, coef = actor4),
#             prior(normal(0, 1.5), class = b, coef = actor5),
#             prior(normal(0, 1.5), class = b, coef = actor6),
#             prior(normal(0, 1.5), class = b, coef = actor7))


mfile <- "models/chimp_m11.4"
# file.remove(paste0(mfile, ".rds"))

get_prior(data = d, family = bernoulli,
          formula = pulled_left ~ 0 + actor + treatment)

priors <- c(prior(normal(0, 0.5), class = b))

bm_11.4 <- brm(data = d, family = bernoulli,
          formula = pulled_left ~ 1 + actor + treatment,
          sample_prior = T, prior = priors,
          cores = 4, chains = 4,
          file = mfile)

```

```{r}

d %>%
  data_grid(actor, treatment) %>%
  add_fitted_draws(bm_11.4) %>%
  median_qi(.value) %>%
  ggplot(aes(x = .value, y = treatment)) + geom_pointintervalh() + facet_wrap(actor~.)

d %>%
  data_grid(actor, treatment) %>%
  add_predicted_draws(bm_11.4) %>%
  median_qi(.prediction) %>%
  ggplot(aes(x = .prediction, y = treatment)) + geom_pointintervalh() + facet_wrap(actor~.)


```

```{r}

mfile <- "models/chimp_m11.5"
# file.remove(paste0(mfile, ".rds"))

get_prior(data = d, family = bernoulli,
          formula = pulled_left ~ 1 + (1|actor) + (1|treatment))

bm_11.5 <- brm(data = d, family = bernoulli,
          formula = pulled_left ~ 1 + (1|actor) + (1|treatment),
          sample_prior = T, #prior = priors,
          cores = 4, chains = 4,
          control = list(adapt_delta = 0.95),
          file = mfile)


```



```{r}

bm_11.5 %>%
  spread_draws(b_Intercept, r_actor[actor, ], r_treatment[treatment, ]) %>%
  median_qi(preds = inv_logit_scaled(b_Intercept + r_actor + r_treatment)) %>%
  ggplot(aes(x = preds, y = treatment)) + geom_pointintervalh() + facet_wrap(actor~.)

```

```{r, fig.width=12}
p1 <- d %>%
        data_grid(actor, treatment) %>%
        add_fitted_draws(bm_11.4) %>%
        median_qi(.value) %>%
        ggplot(aes(x = .value, y = treatment)) + geom_pointintervalh() + facet_wrap(actor~.) +
          ggtitle("Fixed Effects model")

p2 <- d %>%
        data_grid(actor, treatment) %>%
        add_fitted_draws(bm_11.5) %>%
        median_qi(.value) %>%
        ggplot(aes(x = .value, y = treatment)) + geom_pointintervalh() + facet_wrap(actor~.) +
          ggtitle("Random effects model")

grid.arrange(p1, p2, nrow = 1)
```


Actor only model
```{r}

mfile <- "models/chimp_m11.2a"

# file.remove(paste0(mfile, ".rds"))

# d$actor <- as.factor(d$actor)

get_prior(data = d, family = bernoulli,
          # formula = pulled_left | trials(1) ~ 1 + actor,
          formula = pulled_left ~ 1 + actor)

priors <- c(prior(normal(0, 1.5), class = Intercept),
            prior(normal(0, 0.5), class = b))

bm_11.2a <- brm(data = d, family = bernoulli,
          # formula = pulled_left | trials(1) ~ 1 + actor,
          formula = pulled_left ~ 1 + actor,
          sample_prior = T, prior = priors,
          cores = 4, chains = 4,
          file = mfile)

```

pull left behavior of different actors

```{r}

d %>%
  select(actor, pulled_left) %>%
  ggplot(aes(y = pulled_left, x = factor(actor))) + geom_boxplot()


```


```{r}

bm_11.2a$formula
bm_11.2a$family

prior_samples(bm_11.2a)
posterior_samples(bm_11.2a)


d %>%
  data_grid(actor) %>%
  add_predicted_draws(bm_11.2a) %>%
  # filter(actor == 2) %>%
  ggplot(aes(x = .prediction, group = actor, col = factor(actor))) + geom_density()

d %>%
  data_grid(actor) %>%
  add_fitted_draws(bm_11.2a) %>%
  # mutate(.value = inv_logit_scaled(.value)) %>%
  ggplot(aes(x = .value, group = actor, col = factor(actor))) + geom_density()


bm_11.2a%>%
  spread_draws()

```

Rethinking model
```{r}

dat_list <- list(
  pulled_left = d$pulled_left,
  actor = d$actor,
  treatment = as.integer(d$treatment)
)

m11.4 <- ulam(
  alist(
    pulled_left ~ dbinom( 1 , p ) ,
    logit(p) <- a[actor] + b[treatment] ,
    a[actor] ~ dnorm( 0 , 1.5 ),
    b[treatment] ~ dnorm( 0 , 0.5 )
  ) , data=dat_list , chains=4 )

```


```{r}

plot(precis(as.data.frame(inv_logit(extract.samples(m11.4)$a))), xlim = c(0,1))

```

```{r}

m11.5 <- ulam(
  alist(
    pulled_left ~ dbinom( 1 , p ) ,
    logit(p) <- a[actor],
    a[actor] ~ dnorm( 0 , 1.5 )
  ) , data=dat_list , chains=4 )

```

```{r}

precis(as.data.frame(inv_logit(extract.samples(m11.5)$a)))

```
