```{r, include=F, echo=F}

source("../init_settings.R")

library(rethinking)
data(chimpanzees)
d <- chimpanzees

rm(chimpanzees)
detach(package:rethinking, unload = T)
library(brms)

```

Dataset information
* actor : name of actor
* recipient : name of recipient (NA for partner absent condition)
* condition : partner absent (0), partner present (1)
* block : block of trials (each actor x each recipient 1 time)
* trial : trial number (by chimp = ordinal sequence of trials for each chimp, ranges from 1-72; partner present trials were interspersed with partner absent trials)
* prosocial_left : 1 if prosocial (1/1) option was on left; 0 === 2 food items present on right side & 1 === prosocail option present on left
* chose_prosoc : choice chimp made (0 = 1/0 option, 1 = 1/1 option)
* pulled_left : which side did chimp pull (1 = left, 0 = right)

We are trying to study if the chimpanzee pulled the left level when food as present and a pro-social option existed.

There are four combinations:
(1) prosoc_left= 0 and condition= 0: Two food items on right and no partner.
(2) prosoc_left= 1 and condition= 0: Two food items on left and no partner.
(3) prosoc_left= 0 and condition= 1: Two food items on right and partner present.
(4) prosoc_left= 1 and condition= 1: Two food items on left and partner presen

```{r}

d$treatment <- 1 + d$prosoc_left + 2 * d$condition
d$treatment <- as.factor(d$treatment)
d$pulled_left <- as.factor(d$pulled_left)
d$actor <- as.factor(d$actor)
d$block <- as.factor(d$block)

d %>%
  mutate(pulled_left = as.integer(pulled_left) - 1) %>%
  group_by(actor, treatment) %>%
  summarise(pulled_left_frac = sum(pulled_left),
            n())

d %>% mutate(pulled_left = as.integer(pulled_left))

d %>% filter(actor == 1, treatment == 1) %>% summarise(sum(as.integer(pulled_left)))

d %>%
  group_by(actor, treatment) %>%
  summarise(pulled_left_frac = sum(as.integer(pulled_left) - 1)/n()) %>%
    ggplot(aes(x = pulled_left_frac, y = treatment)) + geom_point() + facet_wrap(actor~.) + ggtitle("actor ~ treatment")

d %>%
  group_by(actor, treatment, block) %>%
  summarise(pulled_left_frac = sum(as.integer(pulled_left) - 1)/n()) %>%
  ggplot(aes(x = pulled_left_frac, y = treatment)) +
    geom_point() +
    facet_grid(block ~ actor) + 
    ggtitle("block(y) ~ actor(x)") +
    theme(axis.text.x = element_text(hjust = 1, angle = 45))

```

The question that we will now try to model here is whether the left level was pulled based on the treatment and the actor.

L ~ Binomial(1, p)
p ~ actor[i] = treatment[i]

Setting the priors for the model
```{r}

get_prior(data = d, formula = pulled_left | trials(1) ~ 1, family =binomial)

```

An intercept only simple binomial model

```{r}

mfile <- "models/chimp_bm_test1"

# file.remove(paste0(mfile, ".rds"))

priors <- prior(normal(0,10), class = Intercept)

bm_test1 <- brm(data = d, family = binomial,
          formula = pulled_left | trials(1) ~ 1,
          sample_prior = T, prior = priors,
          cores = 4, chains = 4,
          file = mfile)

mfile <- "models/chimp_bm_test2"

# file.remove(paste0(mfile, ".rds"))

priors <- prior(normal(0,1.5), class = Intercept)

bm_test2 <- brm(data = d, family = binomial,
          formula = pulled_left | trials(1) ~ 1,
          sample_prior = T, prior = priors,
          cores = 4, chains = 4,
          file = mfile)

```

Priot to building the model, what does the model think is plausible?
```{r}

p1 <- prior_samples(bm_test1) %>% 
        inv_logit_scaled() %>%
        ggplot(aes(x = Intercept)) + geom_density()

p2 <- prior_samples(bm_test2) %>% 
        inv_logit_scaled() %>%
        ggplot(aes(x = Intercept)) + geom_density()

grid.arrange(p1, p2, nrow = 1)
```

Adding the treatment column

Identifying the priors
```{r}

get_prior(family = binomial, data = d, formula = pulled_left | trials(1) ~ 1 + treatment)

```


```{r}

mfile <- "models/chimp_m11.2"

# file.remove(paste0(mfile, ".rds"))

get_prior(data = d, family = binomial,
          formula = pulled_left | trials(1) ~ 1 + treatment)

priors <- c(prior(normal(0, 1.5), class = Intercept),
            prior(normal(0, 10), class = b))

bm_11.2 <- brm(data = d, family = bernoulli,
          formula = pulled_left ~ 1 + treatment,
          sample_prior = T, prior = priors,
          cores = 4, chains = 4,
          file = mfile)


mfile <- "models/chimp_m11.3"
# file.remove(paste0(mfile, ".rds"))

priors <- c(prior(normal(0, 1.5), class = Intercept),
            prior(normal(0, 0.5), class = b))

bm_11.3 <- brm(data = d, family = bernoulli,
          formula = pulled_left ~ 1 + treatment,
          sample_prior = T, prior = priors,
          cores = 4, chains = 4,
          file = mfile)

```

```{r}

p1 <- prior_samples(bm_11.2) %>%
        mutate(preds = inv_logit_scaled(b + Intercept)) %>%
        ggplot(aes(x = preds)) + geom_density()
  
p2 <- prior_samples(bm_11.3) %>%
        mutate(preds = inv_logit_scaled(b + Intercept)) %>%
        ggplot(aes(x = preds)) + geom_density()

grid.arrange(p1, p2, nrow =1)

```


```{r}

d %>%
  data_grid(treatment) %>%
  add_fitted_draws(bm_11.2) %>%
  median_qi(.value) %>%
  ggplot(aes(x = .value, y = treatment, xmin = .lower, xmax = .upper)) + geom_pointintervalh()

```

Considering the behavior of the actor only

```{r}

mfile <- "models/bm_a1"
# file.remove(paste0(mfile, ".rds"))

get_prior(family = bernoulli, data = d,
             formula = pulled_left ~ 1 + actor)

priors <- c(prior(normal(0, 1.5), class = Intercept),
            prior(normal(0, 10.5), class = b))

bm_a1 <- brm(family = bernoulli, data = d,
             formula = pulled_left ~ 1 + actor,
             sample_prior = T, prior = priors,
             cores = 4, chains = 4,
             control = list(max_treedepth = 15),
             file = mfile)


```

```{r}

prior_samples(bm_a1) %>%
  inv_logit_scaled() %>%
  ggplot(aes(x = Intercept)) + geom_density()

```

```{r}

d %>%
  data_grid(actor) %>%
  add_fitted_draws(bm_a1) %>%
  median_qi(.value) %>%
  ggplot(aes(x = .value, y = actor, xmin = .lower, xmax = .upper)) + geom_pointintervalh()


```


```{r}

mfile <- "models/bm_a1_h"
# file.remove(paste0(mfile, ".rds"))

get_prior(family = bernoulli, data = d,
             formula = pulled_left ~ 1 + (1|actor))

priors <- c(prior(normal(0, 1.5), class = Intercept))

bm_a1_h <- brm(family = bernoulli, data = d,
             formula = pulled_left ~ 1 + (1|actor),
             sample_prior = T, prior = priors,
             cores = 4, chains = 4,
             control = list(max_treedepth = 15),
             file = mfile)


```

```{r}

p1 <- d %>%
  data_grid(actor) %>%
  add_fitted_draws(bm_a1) %>%
  median_qi(.value) %>%
  ggplot(aes(x = .value, y = actor, xmin = .lower, xmax = .upper)) + geom_pointintervalh() + ggtitle("Fixed Effects Model")

p2 <- d %>%
  data_grid(actor) %>%
  add_fitted_draws(bm_a1_h) %>%
  median_qi(.value) %>%
  ggplot(aes(x = .value, y = actor, xmin = .lower, xmax = .upper)) + geom_pointintervalh() + ggtitle("Random Effects Model")

grid.arrange(p1, p2, nrow = 1)

```

```{r, fig.width=12}

d1 <- d %>%
  data_grid(actor) %>%
  add_fitted_draws(bm_a1) %>%
  ungroup() %>%
  select(actor, .value) %>%
  as_tibble() %>%
  mutate(type = "fixed")

d2 <- d %>%
  data_grid(actor) %>%
  add_fitted_draws(bm_a1_h) %>%
  ungroup() %>%
  select(actor, .value) %>%
  as_tibble() %>%
  mutate(type = "random")

d3 <- rbind(d1, d2)

d3 %>%
  ggplot(aes(x = .value, col = type)) + geom_density(alpha=0.2, aes(fill = type)) + facet_wrap(actor~., scales = "free")

```

Considering both actor and treatment

```{r}


# priors <- c(prior(normal(0, 0.5), class = b, coef = treatment2),
#              prior(normal(0, 0.5), class = b, coef = treatment3),
#              prior(normal(0, 0.5), class = b, coef = treatment4),
#              prior(normal(0, 1.5), class = b, coef = actor1),
#             prior(normal(0, 1.5), class = b, coef = actor2),
#             prior(normal(0, 1.5), class = b, coef = actor3),
#             prior(normal(0, 1.5), class = b, coef = actor4),
#             prior(normal(0, 1.5), class = b, coef = actor5),
#             prior(normal(0, 1.5), class = b, coef = actor6),
#             prior(normal(0, 1.5), class = b, coef = actor7))


mfile <- "models/chimp_m11.4"
# file.remove(paste0(mfile, ".rds"))

get_prior(data = d, family = bernoulli,
          formula = pulled_left ~ 0 + actor + treatment)

priors <- c(prior(normal(0, 0.5), class = b))

bm_11.4 <- brm(data = d, family = bernoulli,
          formula = pulled_left ~ 1 + actor + treatment,
          sample_prior = T, prior = priors,
          cores = 4, chains = 4,
          file = mfile)

```

```{r}

d %>%
  data_grid(actor, treatment) %>%
  add_fitted_draws(bm_11.4) %>%
  median_qi(.value) %>%
  ggplot(aes(x = .value, y = treatment)) + geom_pointintervalh() + facet_wrap(actor~.)

d %>%
  data_grid(actor, treatment) %>%
  add_predicted_draws(bm_11.4) %>%
  median_qi(.prediction) %>%
  ggplot(aes(x = .prediction, y = treatment)) + geom_pointintervalh() + facet_wrap(actor~.)


```

```{r}

mfile <- "models/chimp_m11.5"
# file.remove(paste0(mfile, ".rds"))

get_prior(data = d, family = bernoulli,
          formula = pulled_left ~ 1 + (1|actor) + (1|treatment))

bm_11.5 <- brm(data = d, family = bernoulli,
          formula = pulled_left ~ 1 + (1|actor) + (1|treatment),
          sample_prior = T, #prior = priors,
          cores = 4, chains = 4,
          control = list(adapt_delta = 0.95),
          file = mfile)


```

```{r}

bm_11.5 %>%
  spread_draws(b_Intercept, r_actor[actor, ], r_treatment[treatment, ]) %>%
  median_qi(preds = inv_logit_scaled(b_Intercept + r_actor + r_treatment)) %>%
  ggplot(aes(x = preds, y = treatment)) + geom_pointintervalh() + facet_wrap(actor~.)

```

```{r, fig.width=12}

p1 <- d %>%
        data_grid(actor, treatment) %>%
        add_fitted_draws(bm_11.4) %>%
        median_qi(.value) %>%
        ggplot(aes(x = .value, y = treatment)) + geom_pointintervalh() + facet_wrap(actor~.) +
          ggtitle("Fixed Effects model")

p2 <- d %>%
        data_grid(actor, treatment) %>%
        add_fitted_draws(bm_11.5) %>%
        median_qi(.value) %>%
        ggplot(aes(x = .value, y = treatment)) + geom_pointintervalh() + facet_wrap(actor~.) +
          ggtitle("Random effects model")

grid.arrange(p1, p2, nrow = 1)

```


Actor only model
```{r}

mfile <- "models/chimp_m11.2a"

# file.remove(paste0(mfile, ".rds"))

# d$actor <- as.factor(d$actor)

get_prior(data = d, family = bernoulli,
          # formula = pulled_left | trials(1) ~ 1 + actor,
          formula = pulled_left ~ 1 + actor)

priors <- c(prior(normal(0, 1.5), class = Intercept),
            prior(normal(0, 0.5), class = b))

bm_11.2a <- brm(data = d, family = bernoulli,
          # formula = pulled_left | trials(1) ~ 1 + actor,
          formula = pulled_left ~ 1 + actor,
          sample_prior = T, prior = priors,
          cores = 4, chains = 4,
          file = mfile)

```

pull left behavior of different actors


```{r}

bm_11.2a$formula
bm_11.2a$family

prior_samples(bm_11.2a)
posterior_samples(bm_11.2a)


d %>%
  data_grid(actor) %>%
  add_predicted_draws(bm_11.2a) %>%
  # filter(actor == 2) %>%
  ggplot(aes(x = .prediction, group = actor, col = factor(actor))) + geom_density()

d %>%
  data_grid(actor) %>%
  add_fitted_draws(bm_11.2a) %>%
  # mutate(.value = inv_logit_scaled(.value)) %>%
  ggplot(aes(x = .value, group = actor, col = factor(actor))) + geom_density()

```

Studying contrasts

```{r}

get_variables(bm_11.4)

bm_11.4 %>%
  spread_draws(b_Intercept, b_treatment2, b_treatment3, b_treatment4) %>%
  mutate(db13 = b_Intercept - b_treatment3,
         db24 = b_treatment2 - b_treatment4) %>%
  select(db13, db24) %>%
  summary()

```




