---
title: "R Notebook"
output: html_notebook
---

```{r, echo=F, include=FALSE}

source("../init_settings.R")

library(rethinking)
data("UCBAdmissions")
d <- UCBAdmissions
d <- as.data.frame(d)

rm(UCBAdmissions)
detach(package:rethinking, unload = T)
library(brms)

```

```{r}

d %>%
  pivot_wider(names_from = c(Admit, Gender), values_from = Freq) %>%
  mutate(perc_admit_male = Admitted_Male / (Admitted_Male + Rejected_Male),
         perc_admit_female = Admitted_Female / (Admitted_Female + Rejected_Female),
         perc_admit_overall = (Admitted_Male + Admitted_Female) / (Admitted_Male + Rejected_Male + Admitted_Female + Rejected_Female),
         perc_male_admit = Admitted_Male / (Admitted_Male + Admitted_Female)) %>%
  mutate_at(vars(starts_with("perc_")), ~round(., digits = 2))

d
```

Checking if there is a bias overall

```{r}

d_m1 <- d %>%
          pivot_wider(names_from = c(Admit), values_from = Freq) %>%
          mutate(applications = Admitted + Rejected)
d_m1
```

```{r}

mfile <- "models/gsa_m1_defp"
# file.remove(paste0(mfile, ".rds"))

# get_prior(data = d_m1, family = binomial, prior = priors,
#           formula = Admitted | trials(applications) ~ 1 + Gender)

gsa_m1_defp <- brm(data = d_m1, family = binomial,
              formula = Admitted | trials(applications) ~ 1 + Gender,
              chains = 4, cores = 4,
              sample_prior = T,
              file = mfile)

```

mcmc traces
```{r}

mcmc_trace(gsa_m1_defp)
mcmc_dens(gsa_m1_defp)

```


Prior simulations
```{r}

prior_samples(gsa_m1_defp) %>%
  inv_logit_scaled() %>%
  ggplot(aes(x = Intercept)) + geom_density()

```

Posterior predictive checks

```{r}

d_m1 %>%
  data_grid(Gender, applications) %>%
  add_predicted_draws(gsa_m1_defp) %>%
  median_qi() %>%
  ggplot(aes(x = .prediction, y = applications, xmax = .upper, xmin = .lower, col = Gender)) + geom_pointintervalh()

gsa_m1_defp %>%
  spread_draws(b_Intercept, b_GenderFemale) %>%
  transmute(Male_prob = b_Intercept,
            Female_prob = b_Intercept + b_GenderFemale) %>%
  inv_logit_scaled() %>%
  gather(prob_types, values) %>%
  ggplot(aes(x = values, col = prob_types)) + geom_density()

```



```{r}
mfile <- "models/gsa_m1_regp"
# file.remove(paste0(mfile, ".rds"))

# get_prior(data = d_m1, family = binomial, prior = priors,
#           formula = Admitted | trials(applications) ~ 1 + Gender)

# priors <- c(prior(normal(0,1.5), class = b, coef = "GenderFemale"),
#             prior(normal(0,1.5), class = Intercept))

priors <- c(prior(normal(0,1.5), class = Intercept))


gsa_m1_regp <- brm(data = d_m1, family = binomial,
              formula = Admitted | trials(applications) ~ 1 + Gender,
              chains = 4, cores = 4,
              sample_prior = T, prior = priors,
              file = mfile)

```


Prior predictive simulations

```{r}

prior_samples(gsa_m1_regp) %>%
  inv_logit_scaled() %>%
  ggplot(aes(x = Intercept)) + geom_density()
  # gather(ptypes, pvals ) %>%
  # ggplot(aes(x = pvals, col = ptypes)) + geom_density()

d_m1 %>%
  data_grid(Gender, applications) %>%
  add_predicted_draws(gsa_m1_regp) %>%
  ggplot(aes(x = factor(applications), y = .prediction, fill = Gender)) + geom_boxplot()

```

Comparing prediction of regularized priors and default priors

```{r}

regp <- d_m1 %>%
          data_grid(Gender, applications) %>%
          add_predicted_draws(gsa_m1_regp) %>%
          mutate(itype = "regp")

defp <- d_m1 %>%
          data_grid(Gender, applications) %>%
          add_predicted_draws(gsa_m1_defp) %>%
          mutate(itype = "defp")

preds <- regp %>% rbind(defp)

preds %>%
  ggplot(aes(x = factor(applications), y = .prediction, col = itype)) +
    geom_boxplot() +
    facet_grid(Gender~., scales = "free")
```

Density based comparison
```{r}

defp <- gsa_m1_defp %>%
  spread_draws(b_Intercept, b_GenderFemale) %>%
  transmute(Male_prob = b_Intercept,
            Female_prob = b_Intercept + b_GenderFemale) %>%
  inv_logit_scaled() %>%
  mutate(itype = "defp")

regp <- gsa_m1_regp %>%
  spread_draws(b_Intercept, b_GenderFemale) %>%
  transmute(Male_prob = b_Intercept,
            Female_prob = b_Intercept + b_GenderFemale) %>%
  inv_logit_scaled() %>%
  mutate(itype = "regp")

preds <- defp %>% rbind(regp)

preds %>%
  gather(gender, probs, -itype) %>%
  ggplot(aes(x = probs, col = gender, fill = itype)) + geom_density(alpha = 0.2)

```


Regularizartion doesn't seem to have an impact on the predictions

Now analyzing the acceptance rate per group

```{r}



```

```{r}

mfile <- "models/gsa_m2_defp"
# file.remove(paste0(mfile, ".rds"))

# get_prior(data = d_m1, family = binomial, prior = priors,
#           formula = Admitted | trials(applications) ~ 1 + Gender + Dept)

gsa_m2_defp <- brm(data = d_m1, family = binomial,
              formula = Admitted | trials(applications) ~ 1 + Gender + Dept,
              chains = 4, cores = 4,
              sample_prior = T,
              file = mfile)

```

Checking the priors in the model
```{r}

prior_samples(gsa_m2_defp) %>%
  inv_logit_scaled() %>%
  ggplot(aes(x = Intercept)) + geom_density()

```


```{r}

d_m1 %>%
  data_grid(Gender, Dept, applications) %>%
  add_predicted_draws(gsa_m2_defp) %>%
  median_qi(.prediction) %>%
  ggplot(aes(x = factor(applications), y = .prediction, ymin = .lower, ymax = .upper, col = Gender)) +
    geom_pointinterval() +
    facet_grid(Dept~., scales = "free")

d_m1 %>%
  data_grid(Gender, Dept) %>%
  mutate(applications = 100) %>%
  add_predicted_draws(gsa_m2_defp) %>%
  ggplot(aes(x = .prediction, fill = Gender, col = Gender)) +
    geom_density(alpha = 0.2) +
    facet_wrap(Dept~., scales = "free")

```

Using regularising priors

```{r}

mfile <- "models/gsa_m2_regp"
# file.remove(paste0(mfile, ".rds"))

# get_prior(data = d_m1, family = binomial, prior = priors,
#           formula = Admitted | trials(applications) ~ 1 + Gender + Dept)

priors <- c(prior(normal(0, 1.5), class = Intercept),
            prior(normal(0, 0.5), class = b))

gsa_m2_regp <- brm(data = d_m1, family = binomial,
              formula = Admitted | trials(applications) ~ 1 + Gender + Dept,
              chains = 4, cores = 4,
              prior = priors, sample_prior = T,
              file = mfile)

```

mcmc chains
```{r}

mcmc_trace(gsa_m2_regp)
mcmc_dens(gsa_m2_regp)

```


Prior simulation
```{r}

gsa_m2_regp %>%
  prior_samples() %>%
  transmute(prior_preds = Intercept + b) %>%
  inv_logit_scaled() %>%
  ggplot(aes(x = prior_preds)) + geom_density()

```

```{r}

defp <- d_m1 %>%
  data_grid(Gender, Dept) %>%
  mutate(applications = 100) %>%
  add_predicted_draws(gsa_m2_defp) %>%
  ungroup() %>%
  select(Gender, Dept, .prediction) %>%
  mutate(itype = "defp")

regp <- d_m1 %>%
  data_grid(Gender, Dept) %>%
  mutate(applications = 100) %>%
  add_predicted_draws(gsa_m2_regp) %>%
  ungroup() %>%
  select(Gender, Dept, .prediction, -.row) %>%
  mutate(itype = "regp")
  

preds <- rbind(defp, regp)
preds

preds %>%
  ggplot(aes(x = .prediction, fill = itype, col = itype)) +
    geom_density(alpha = 0.2) +
    facet_wrap(Dept~Gender, scales = "free")  

d_m1 %>%
  data_grid(Gender, Dept) %>%
  mutate(applications = 100) %>%
  add_predicted_draws(gsa_m2_regp) %>%
  ggplot(aes(x = .prediction, fill = Gender, col = Gender)) +
    geom_density(alpha = 0.2) +
    facet_wrap(Dept~., scales = "free")

```

Trying out hierarchical models

```{r}

mfile <- "models/gsa_m3_defp"
# file.remove(paste0(mfile, ".rds"))

# get_prior(data = d_m1, family = binomial, prior = priors,
#           formula = Admitted | trials(applications) ~ 1 + (1|Gender) + (1|Dept))

gsa_m3_defp <- brm(data = d_m1, family = binomial,
              formula = Admitted | trials(applications) ~ 1 + (1|Gender) + (1|Dept),
              chains = 4, cores = 4,
              sample_prior = T,
              control = list(adapt_delta = 0.99,
                             max_treedepth = 15),
              file = mfile)


```

Examining the chains

```{r}

mcmc_trace(gsa_m3_defp)

```


```{r}

d_m1 %>%
  data_grid(Gender, Dept) %>%
  mutate(applications = 100) %>%
  add_predicted_draws(gsa_m3_defp) %>%
  ggplot(aes(x = .prediction, fill = Gender, col = Gender)) +
    geom_density(alpha = 0.2) +
    facet_wrap(Dept~., scales = "free")

```

Using regularized priors

```{r}

mfile <- "models/gsa_m3_regp"
# file.remove(paste0(mfile, ".rds"))

# get_prior(data = d_m1, family = binomial, prior = priors,
#           formula = Admitted | trials(applications) ~ 1 + (1|Gender) + (1|Dept))

priors <- c(prior(normal(0, 1.5), class = Intercept))

gsa_m3_regp <- brm(data = d_m1, family = binomial,
              formula = Admitted | trials(applications) ~ 1 + (1|Gender) + (1|Dept),
              chains = 4, cores = 4,
              sample_prior = T, prior = priors,
              control = list(adapt_delta = 0.99,
                             max_treedepth = 15),
              file = mfile)

```

