---
title: "R Notebook"
output: html_notebook
---

```{r, echo=F, include=F}

source("../init_settings.R")

```

Reading the data set

```{r, echo=F}

library(rethinking)
data(WaffleDivorce)
d <- WaffleDivorce

rm(WaffleDivorce)
detach(package:rethinking, unload = T)
library(brms)
library(tidyverse)

```

```{r}

mu_age <- mean(d$MedianAgeMarriage)
sd_age <- sd(d$MedianAgeMarriage)

mu_d <- mean(d$Divorce)
sd_d <- sd(d$Divorce)

mu_mr <- mean(d$Marriage)
sd_mr <- sd(d$Marriage)

d <- d %>% mutate(MedianAgeMarriage_s = (MedianAgeMarriage - mu_age)/ sd_age,
                  Divorce_s = (Divorce - mu_d) / sd_d,
                  Marriage_s = (Marriage - mu_mr) / sd_mr)

d

```

# Impact of age on divorce rate

## Simple visualization
```{r}

d %>%
  ggplot(aes(x = MedianAgeMarriage_s, y = Divorce_s)) +
    geom_point() +
    geom_smooth(method = "lm")

```

Lower the age of marraige, higher the divorce rate

```{r}

mfile <- "models/mra_m1"
# file.remove(paste0(mfile, ".rds"))

# get_prior(family = gaussian, data = d,
#           formula = Divorce_s ~ 1 + MedianAgeMarriage_s)

priors <- c(prior(normal(0, 0.5), class = b, coef = "MedianAgeMarriage_s"),
            prior(normal(0, 0.5), class = Intercept))

mra_m1 <- brm(family = gaussian, data = d,
          formula = Divorce_s ~ 1 + MedianAgeMarriage_s,
          cores = 4, chains = 4,
          sample_prior = T, prior = priors,
          control = list(max_treedepth = 15),
          file = mfile)

```

Prior simulations
```{r}

mcmc_dens(mra_m1)
mcmc_trace(mra_m1)
stanplot(mra_m1)

```

```{r}

mra_m1 %>%
  log_lik() %>%
  as_tibble(.name_repair = "unique") %>%
  transmute(lp = rowSums(.)) %>%
  ggplot(aes(x = lp)) + geom_density()

```
Prior simulations
```{r}

posterior_samples(mra_m1) %>%
  select(prior_Intercept, b_Intercept) %>%
  pivot_longer(names_to = "Intercept_type", 
               values_to = "Intercept_values", 
               cols = c(b_Intercept, prior_Intercept)) %>%
  ggplot(aes(x = Intercept_values, col = Intercept_type)) +
    geom_density(aes(fill = Intercept_type), alpha = 0.2) +
    ggtitle("Compare Intercepts")

posterior_samples(mra_m1) %>%
  select(prior_b, b_MedianAgeMarriage_s) %>%
  pivot_longer(names_to = "slope_type", 
               values_to = "slope_values", 
               cols = c(prior_b, b_MedianAgeMarriage_s)) %>%
  ggplot(aes(x = slope_values, col = slope_type)) +
    geom_density(aes(fill = slope_type), alpha = 0.2) +
    ggtitle("Compare Slopes")

```

Posterior checks
```{r}
pp_check(mra_m1, type = "dens_overlay")
pp_check(mra_m1, type = "error_hist")
pp_check(mra_m1, type = "boxplot")
pp_check(mra_m1, type = "ribbon")
```

```{r}

d %>%
  select(MedianAgeMarriage, MedianAgeMarriage_s, Divorce_s, Divorce) %>%
  add_fitted_draws(mra_m1) %>%
  mutate(preds = .value * sd_d + mu_d) %>%
  ggplot(aes(y = preds, x = MedianAgeMarriage)) + geom_point() +
    geom_point(aes(y = Divorce, x = MedianAgeMarriage), color = "lightblue")

d %>%
  select(MedianAgeMarriage, MedianAgeMarriage_s, Divorce_s, Divorce) %>%
  add_predicted_draws(mra_m1) %>%
  mutate(preds = .prediction * sd_d + mu_d) %>%
  ggplot() +
    geom_point(aes(y = preds, x = MedianAgeMarriage)) +
    geom_point(aes(y = Divorce, x = MedianAgeMarriage), color = "lightblue")

```


# Impact of marriage rate on divorce rate

## Simple visualization
```{r}

d %>%
  ggplot(aes(x = Marriage_s, y = Divorce_s)) +
    geom_point() +
    geom_smooth(method = "lm")

```

Higher the marriage rate, higher the divorce rate

```{r}

mfile <- "models/mra_m2"
# file.remove(paste0(mfile, ".rds"))

# get_prior(family = gaussian, data = d,
#           formula = Divorce_s ~ 1 + MedianAgeMarriage_s)

priors <- c(prior(normal(0, 0.5), class = b, coef = "Marriage_s"),
            prior(normal(0, 0.5), class = Intercept))

mra_m2 <- brm(family = gaussian, data = d,
          formula = Divorce_s ~ 1 + Marriage_s,
          cores = 4, chains = 4,
          sample_prior = T, prior = priors,
          control = list(max_treedepth = 15),
          file = mfile)

```

Prior simulations
```{r}

mcmc_dens(mra_m2)
mcmc_trace(mra_m2)
stanplot(mra_m2)

```

```{r}

mra_m2 %>%
  log_lik() %>%
  as_tibble(.name_repair = "unique") %>%
  transmute(lp = rowSums(.)) %>%
  ggplot(aes(x = lp)) + geom_density()

```
Prior simulations
```{r}

posterior_samples(mra_m2) %>%
  select(prior_Intercept, b_Intercept) %>%
  pivot_longer(names_to = "Intercept_type", 
               values_to = "Intercept_values", 
               cols = c(b_Intercept, prior_Intercept)) %>%
  ggplot(aes(x = Intercept_values, col = Intercept_type)) +
    geom_density(aes(fill = Intercept_type), alpha = 0.2) +
    ggtitle("Compare Intercepts")

posterior_samples(mra_m2) %>%
  select(prior_b, b_Marriage_s) %>%
  pivot_longer(names_to = "slope_type", 
               values_to = "slope_values", 
               cols = c(prior_b, b_Marriage_s)) %>%
  ggplot(aes(x = slope_values, col = slope_type)) +
    geom_density(aes(fill = slope_type), alpha = 0.2) +
    ggtitle("Compare Slopes")

```

Posterior checks
```{r}
pp_check(mra_m2, type = "dens_overlay")
pp_check(mra_m2, type = "error_hist")
pp_check(mra_m2, type = "boxplot")
pp_check(mra_m2, type = "ribbon")
```

```{r}

d %>%
  select(Marriage, Marriage_s, Divorce_s, Divorce) %>%
  add_fitted_draws(mra_m2) %>%
  mutate(preds = .value * sd_d + mu_d) %>%
  ggplot(aes(y = preds, x = Marriage)) + geom_point() +
    geom_point(aes(y = Divorce, x = Marriage), color = "lightblue")

d %>%
  select(Marriage, Marriage_s, Divorce_s, Divorce) %>%
  add_predicted_draws(mra_m2) %>%
  mutate(preds = .prediction * sd_d + mu_d) %>%
  ggplot() +
    geom_point(aes(y = preds, x = Marriage)) +
    geom_point(aes(y = Divorce, x = Marriage), color = "lightblue")

```

Compare the 2 models
```{r}

loo_compare(waic(mra_m1), waic(mra_m2))

```

```{r}

mfile <- "models/mra_m3"
# file.remove(paste0(mfile, ".rds"))

# get_prior(family = gaussian, data = d,
#           formula = Divorce_s ~ 1 + MedianAgeMarriage_s)

priors <- c(prior(normal(0, 0.5), class = b),
            prior(normal(0, 0.5), class = Intercept))

mra_m3 <- brm(family = gaussian, data = d,
          formula = Divorce_s ~ 1 + Marriage_s + MedianAgeMarriage_s,
          cores = 4, chains = 4,
          sample_prior = T, prior = priors,
          control = list(max_treedepth = 15))

```

```{r}
mcmc_dens(mra_m3)
mcmc_trace(mra_m3)
stanplot(mra_m3)

```

```{r}

mra_m3 %>%
  log_lik() %>%
  as_tibble(.name_repair = "unique") %>%
  transmute(lp = rowSums(.)) %>%
  ggplot(aes(x = lp)) + geom_density()

```

```{r}

get_variables(mra_m3)

posterior_samples(mra_m3) %>%
  select(prior_Intercept, b_Intercept) %>%
  pivot_longer(names_to = "Intercept_type", 
               values_to = "Intercept_values", 
               cols = c(b_Intercept, prior_Intercept)) %>%
  ggplot(aes(x = Intercept_values, col = Intercept_type)) +
    geom_density(aes(fill = Intercept_type), alpha = 0.2) +
    ggtitle("Compare Intercepts")

posterior_samples(mra_m3) %>%
  select(prior_b, b_Marriage_s) %>%
  pivot_longer(names_to = "slope_type", 
               values_to = "slope_values", 
               cols = c(prior_b, b_Marriage_s)) %>%
  ggplot(aes(x = slope_values, col = slope_type)) +
    geom_density(aes(fill = slope_type), alpha = 0.2) +
    ggtitle("Compare Slopes - Marraige Rate")

posterior_samples(mra_m3) %>%
  select(prior_b, b_MedianAgeMarriage_s) %>%
  pivot_longer(names_to = "slope_type", 
               values_to = "slope_values", 
               cols = c(prior_b, b_MedianAgeMarriage_s)) %>%
  ggplot(aes(x = slope_values, col = slope_type)) +
    geom_density(aes(fill = slope_type), alpha = 0.2) +
    ggtitle("Compare Slopes - Age")

```

Posterior checks
```{r}
pp_check(mra_m3, type = "dens_overlay")
pp_check(mra_m3, type = "error_hist")
pp_check(mra_m3, type = "boxplot")
pp_check(mra_m3, type = "ribbon")
```

Setting median marriage to 0 and looking at predictions

```{r}

d %>%
  select(Marriage, Marriage_s, Divorce_s, Divorce) %>%
  mutate(MedianAgeMarriage_s = 0,
         MedianAgeMarriage = MedianAgeMarriage_s * sd_age + mu_age) %>%
  add_fitted_draws(mra_m3) %>%
  mutate(preds = .value * sd_d + mu_d) %>%
  ggplot(aes(y = preds, x = Marriage)) + geom_point() +
    geom_point(aes(y = Divorce, x = Marriage), color = "lightblue") +
    ggtitle("Counterfactual (Fitted Draws): setting scaled median age to 0")

d %>%
  select(Marriage, Marriage_s, Divorce_s, Divorce) %>%
  mutate(MedianAgeMarriage_s = 0,
         MedianAgeMarriage = MedianAgeMarriage_s * sd_age + mu_age) %>%
  add_predicted_draws(mra_m3) %>%
  mutate(preds = .prediction * sd_d + mu_d) %>%
  ggplot() +
    geom_point(aes(y = preds, x = Marriage)) +
    geom_point(aes(y = Divorce, x = Marriage), color = "lightblue") +
    ggtitle("Counterfactual (Predicted Draws): setting scaled median age to 0")


```

Setting marriage rate to 0 and looking at predictions

```{r}

d %>%
  select(Divorce_s, Divorce,MedianAgeMarriage_s , MedianAgeMarriage) %>%
  mutate(Marriage_s = 0,
         Marriage = Marriage_s * sd_mr + mu_mr) %>%
  add_fitted_draws(mra_m3) %>%
  mutate(preds = .value * sd_d + mu_d) %>%
  ggplot(aes(y = preds, x = MedianAgeMarriage)) + geom_point() +
    geom_point(aes(y = Divorce, x = MedianAgeMarriage), color = "lightblue") +
    ggtitle("Counterfactual (Fitted Draws): setting scaled marriage rate to 0")

d %>%
  select(Divorce_s, Divorce, MedianAgeMarriage, MedianAgeMarriage_s) %>%
  mutate(Marriage_s = 0,
         Marriage = Marriage_s * sd_mr + mu_mr) %>%
  add_predicted_draws(mra_m3) %>%
  mutate(preds = .prediction * sd_d + mu_d) %>%
  ggplot() +
    geom_point(aes(y = preds, x = MedianAgeMarriage)) +
    geom_point(aes(y = Divorce, x = MedianAgeMarriage), color = "lightblue") +
    ggtitle("Counterfactual (Predicted Draws): setting scaled marriage rate to 0")

```

As we can clearly see, there is a trend observed with changing age and fixed marraige rate, than the reverse.  This is seen from the multivariate regression