Learning from https://www.rensvandeschoot.com/tutorials/generalised-linear-models-with-brms/

```{r}

require(gridExtra)
require(tidyverse)
require(haven)
require(ROCR)
require(brms)
require(modelr)
require(tidybayes)
require(bayesplot)
require(grid)

```

Importing data
```{r}
ThaiEdu_Raw <- read_sav("https://github.com/MultiLevelAnalysis/Datasets-third-edition-Multilevel-book/blob/master/chapter%206/Thaieduc/thaieduc.sav?raw=true")
ThaiEdu_Raw

write.csv2(ThaiEdu_Raw, file = "ThaiEducation.csv")
```

```{r}

df <- read.csv("ThaiEducation.csv", sep = ";")
df

```


```{r}

df <- df %>%
  mutate(SCHOOLID = factor(SCHOOLID),
         SEX = if_else(SEX == 0, "girl", "boy"),
         SEX = factor(SEX, levels = c("girl", "boy")),
         PPED = if_else(PPED == 0, "no", "yes"),
         PPED = factor(PPED, levels = c("no", "yes")))

df <- df %>% filter(!(is.na(MSESC)))

head(df)

```

Question 1: what are the effects of gender and preschool education on whether a pupil repeats a grade?

Simple explorations
```{r}

# p1 <- df %>%
#   group_by(SEX, REPEAT) %>%
#   summarise(count = n()) %>%
#   ggplot(aes(x = SEX, y = count, fill = factor(REPEAT))) +
#     geom_bar(stat = "identity")
# 
# p2 <- df %>%
#   group_by(PPED, REPEAT) %>%
#   summarise(count = n()) %>%
#   ggplot(aes(x = PPED, y = count, fill = factor(REPEAT))) +
#     geom_bar(stat = "identity")
# 
# grid.arrange(p1, p2, nrow = 1)

df %>%
  group_by(SEX, PPED, REPEAT) %>%
  summarise(count = n()) %>%
  gather(key, lvls, -count, -REPEAT) %>%
  ggplot(aes(x = lvls, y = count, fill = factor(REPEAT))) + geom_bar(stat = "identity")

df
```

Simple brms models
```{r}

m1 <- brm(data = df, family = bernoulli(),
          formula = REPEAT ~ SEX + PPED,
          chains = 4, cores = 4)

```

Checking model convergence
```{r}
mcmc_trace(m1)

```

The above plot shows iterations after burn-in.  There sones't seem to be divergences.

Below we check of strong auto correlations that could bias the variance estimates.

```{r}

stanplot(m1, type = "acf_bar")

```
Model interpretation
```{r}

summary(m1)

mcmc_dens(m1)

stanplot(m1, type = "areas", prob = 0.95)

```

Using spread draws
```{r}

get_variables(m1)

m1 %>%
  spread_draws(b_Intercept, b_SEXboy, b_PPEDyes) %>%
  median_qi(preds = exp(b_Intercept + b_SEXboy + b_PPEDyes))

table(interaction(df$SEX, df$PPED))

```

Prediction using the current model

```{r}

df %>%
  data_grid(SEX, PPED) %>%
  add_fitted_draws(m1) %>%
  ggplot(aes(x = .value, y = interaction(SEX, PPED))) +
    stat_pointintervalh() +
    coord_flip() +
    xlab("predicted probability") +
    scale_x_continuous(breaks = seq(0, 0.24, 0.02))

```

Bayesian binomial logistic regression:

data prep

```{r}

ThaiEdu_Prop <- df %>%
  group_by(SCHOOLID, MSESC) %>%
  summarise(REPEAT = sum(REPEAT),
            TOTAL = n()) %>%
  ungroup()

head(ThaiEdu_Prop)

```

```{r}

m2 <- brm(data = ThaiEdu_Prop, family = binomial(link = "logit"),
          formula = REPEAT | trials(TOTAL) ~ MSESC,
          cores = 4, chains = 4)

```

```{r}

mcmc_trace(m2)

```
