---
title: "R Notebook"
output: html_notebook
---

generate training data

```{r, echo=F, include=F}

source("../init_settings.R")

```


```{r}
n_unq <- 50
reps <- 3

df <- tibble(
  x = rep(seq_len(n_unq), reps),
  g1 = rnorm(n_unq * reps, mean = 5, sd = 2) * x + rnorm(n_unq * reps, mean = 0, sd = 0.5),
  g2 = rnorm(n_unq * reps, mean = -15, sd = 2) * x + rnorm(n_unq * reps, mean = 200, sd = 0.5),
  g3 = rnorm(n_unq * reps, mean = 25, sd = 2) * x + rnorm(n_unq * reps, mean = -150, sd = 0.5)
)

mdf <- df %>%
  pivot_longer(cols = c(g1, g2, g3), names_to = "grps", values_to = "y")


mu_x <- mean(mdf$x)
sd_x <- sd(mdf$x)

mu_y <- mean(mdf$y)
sd_y <- sd(mdf$y)

mdf$scaled_x <- (mdf$x - mu_x) / sd_x
mdf$scaled_y <- (mdf$y - mu_y) / sd_y

ggplot(mdf, aes(x = x, y = y, col = grps)) + geom_point() + geom_smooth(method = "lm")

ggplot(mdf, aes(x = x, y = y)) + geom_point() + geom_smooth(method = "lm")

```

Simple regression model
```{r}

summary(lm(y ~ x, data = mdf))

```

Switch to rebuild models

```{r}

rebuild <- F

```



Building a model considering only the numeric covariate.  This model only allows for a fixed intercept and slope.

```{r}

mfile <- "models/vsi_v2_m0"

if (rebuild) {
  file.remove(paste0(mfile, ".rds"))
}

vsi_v2_m0 <- brm(data = mdf, family = gaussian,
              formula = y ~ 1 + x,
              sample_prior = T,
              chains = 4, cores = 4,
              file = mfile)

```

Assessing vsi_v2_m0, the basic diagnostics first
1. The variance associated with the slope is very high
2. The variance parameter is also pretty high ~ 240 - 260


```{r}

summary(vsi_v2_m0)

mcmc_dens(vsi_v2_m0)

```

Computing log pointwise density

```{r}

waic(vsi_v2_m0)

```


```{r}

lppd <- vsi_v2_m0 %>%
  log_lik() %>%
  as_tibble() %>%
  summarise_all(~mean(.)) %>%
  unlist() %>% unname()

pWAIC <- vsi_v2_m0 %>%
  log_lik() %>%
  as_tibble() %>%
  summarise_all(~var(.)) %>%
  unlist() %>% unname()

estimate <- -2*(sum(lppd) - sum(pWAIC))
print(paste0("Estimate: ", round(estimate,2)))
SE <- sqrt(450 * var(-2*(lppd - pWAIC)))
print(paste0("SE :", round(SE,2)))

```



```{r}

mdf %>%
  data_grid(x) %>%
  add_predicted_draws(vsi_v2_m0) %>%
  ggplot(aes(x = x, y = .prediction)) +
    geom_point() +
    ggtitle(paste0("Predicted Draws, ", vsi_v2_m0$formula))

mdf %>%
  data_grid(x) %>%
  add_fitted_draws(vsi_v2_m0) %>%
  ggplot(aes(x = x, y = .value)) +
    geom_point() +
    ggtitle(paste0("Fitted Draws, ", vsi_v2_m0$formula))

ggplot(mdf, aes(x = x, y = y, col = grps)) +
  geom_point() +
  ggtitle("Original Data")

```

PPC checks
```{r}

pp_check(vsi_v2_m0) + ggtitle(vsi_v2_m0$formula)
pp_check(vsi_v2_m0, type = "hist", binwidth = 25) + ggtitle(vsi_v2_m0$formula)
pp_check(vsi_v2_m0, type = "error_hist", binwidth = 25) + ggtitle(vsi_v2_m0$formula)
pp_check(vsi_v2_m0, type = "stat", binwidth = 25) + ggtitle(vsi_v2_m0$formula)
pp_check(vsi_v2_m0, type = "scatter") + ggtitle(vsi_v2_m0$formula)
pp_check(vsi_v2_m0, type = "loo_intervals") + ggtitle(vsi_v2_m0$formula)

```



```{r}

mfile <- "models/vsi_v2_m1"
if (rebuild) {
  file.remove(paste0(mfile, ".rds"))
}

vsi_v2_m1 <- brm(data = mdf, family = gaussian,
              formula = y ~ 1 + x + grps,
              sample_prior = T,
              chains = 4, cores = 4,
              file = mfile)

```

First need a way to figure out that the basic model is bad and it can be consistently improved.

```{r}

summary(vsi_v2_m1)

```

```{r}

mcmc_dens(vsi_v2_m1)

```

Recreating the log probability distribution
```{r}

vsi_v2_m1 %>%
  log_lik() %>%
  as_tibble() %>%
  transmute(rs = rowSums(.)) %>%
  ggplot(aes(x = rs)) +
    geom_density(color = "darkblue", fill = "lightblue") +
    ggtitle("lppd distribution")

```



```{r, fig.width=16, fig.height=6}

pp_check(vsi_v2_m1) + ggtitle(vsi_v2_m1$formula)
pp_check(vsi_v2_m1, type = "hist") + ggtitle(vsi_v2_m1$formula)
pp_check(vsi_v2_m1, type = "error_hist")
pp_check(vsi_v2_m1, type = "stat")
pp_check(vsi_v2_m1, type = "scatter")
pp_check(vsi_v2_m1, type = "intervals_grouped", group = "grps")

```


```{r}

mdf %>%
  data_grid(x, grps) %>%
  add_predicted_draws(vsi_v2_m1) %>%
  ggplot(aes(x = x, y = .prediction, col = grps, position = "jitter")) + geom_jitter(width = 0.25, alpha = 0.01)

```

```{r}

loo_compare(waic(vsi_v2_m0), waic(vsi_v2_m1))

```



```{r, fig.width=12, fig.height=6}
p1 <- stanplot(vsi_v2_m0) + ggtitle(vsi_v2_m0$formula)
p2 <- stanplot(vsi_v2_m1) + ggtitle(vsi_v2_m1$formula)

grid.arrange(p1, p2, nrow = 1)
```


```{r}

mfile <- "models/vsi_v2_m2"
if (rebuild) {
  file.remove(paste0(mfile, ".rds"))
}

vsi_v2_m2 <- brm(data = mdf, family = gaussian,
         formula = y ~ 1 + x + (1|grps),
         sample_prior = T,
         chains = 4, cores = 4,
         file = mfile)

```

```{r}

p1 <- vsi_v2_m2 %>%
  spread_draws(b_Intercept, r_grps[grps, ]) %>%
  median_qi(r_grps) %>%
  ggplot(aes(y = grps, x = r_grps, xmin = .lower, xmax = .upper)) +
    geom_pointintervalh() +
    ggtitle("grp intercepts only")

p2 <- vsi_v2_m2 %>%
  spread_draws(b_Intercept, r_grps[grps, ]) %>%
  mutate(grps_mean = b_Intercept + r_grps) %>%
  median_qi(grps_mean) %>%
  ggplot(aes(y = grps, x = grps_mean, xmin = .lower, xmax = .upper)) +
    geom_pointintervalh() +
    ggtitle("grp intercepts and global intercept only")

grid.arrange(p1, p2, nrow = 1)

```


```{r, fig.width=16, fig.height=6}

pp_check(vsi_v2_m1, type = "intervals_grouped", group = "grps") + ggtitle(vsi_v2_m1$formula)
pp_check(vsi_v2_m2, type = "intervals_grouped", group = "grps") + ggtitle(vsi_v2_m2$formula)

```

```{r}
loo_compare(waic(vsi_v2_m0), waic(vsi_v2_m1), waic(vsi_v2_m2))
```


```{r, fig.width=12, fig.height=6}

p1 <- mdf %>%
  data_grid(x, grps) %>%
  add_predicted_draws(vsi_v2_m2) %>%
  group_by(x, grps) %>%
  sample_n(100) %>%
  ungroup() %>%
  ggplot(aes(x = x, y = .prediction, col = grps)) +
    geom_point() +
    ggtitle(paste0("Predicted Draws ", vsi_v2_m2$formula))

p2 <- mdf %>%
  select(x, grps) %>%
  add_fitted_draws(vsi_v2_m2) %>%
  group_by(x, grps) %>%
  sample_n(100) %>%
  ungroup() %>%
  ggplot(aes(x = x, y = .value, col = grps)) +
    geom_point() +
    ggtitle(paste0("Fitted Draws ", vsi_v2_m2$formula))

grid.arrange(p1, p2, nrow = 1)

```

```{r}

mfile <- "models/vsi_v2_m3"
if (rebuild) {
  file.remove(paste0(mfile, ".rds"))
}

vsi_v2_m3 <- brm(data = mdf, family = gaussian,
            formula = y ~ 1 + (1 + x|grps),
            sample_prior = T,
            cores = 4, chains = 4,
            file = mfile,
            control = list(adapt_delta = 0.99,
                           max_treedepth = 15))

```

```{r}

mfile <- "models/vsi_v2_m4"
if (rebuild) {
  file.remove(paste0(mfile, ".rds"))
}

vsi_v2_m4 <- brm(data = mdf, family = gaussian,
            formula = y ~ 1 + (1 + scaled_x|grps),
            sample_prior = T,
            cores = 4, chains = 4,
            file = mfile,
            control = list(adapt_delta = 0.9999,
                           max_treedepth = 15))

```

```{r}

mfile <- "models/vsi_v2_m5"
if (rebuild) {
  file.remove(paste0(mfile, ".rds"))
}

vsi_v2_m5 <- brm(data = mdf, family = gaussian,
            formula = y ~ 1 + (scaled_x|grps),
            sample_prior = T,
            cores = 4, chains = 4,
            file = mfile,
            control = list(adapt_delta = 0.9999,
                           max_treedepth = 15))

```


```{r}

vsi_v2_m4 %>%
  spread_draws(b_Intercept, r_grps[grps, Intercept]) %>%
  filter(Intercept == "Intercept") %>%
  mutate(b_intcpt = b_Intercept + r_grps) %>%
  median_qi(b_intcpt)

vsi_v2_m5 %>%
  spread_draws(b_Intercept, r_grps[grps, Intercept]) %>%
  filter(Intercept == "Intercept") %>%
  mutate(b_intcpt = b_Intercept + r_grps) %>%
  median_qi(b_intcpt)

get_variables(vsi_v2_m4)
print("\n")
get_variables(vsi_v2_m5)

loo_compare(waic(vsi_v2_m4), waic(vsi_v2_m5))

```


```{r, fig.width=12, fig.height=6}

p1 <- mcmc_dens(vsi_v2_m4) + ggtitle(vsi_v2_m4$formula)
p2 <- mcmc_dens(vsi_v2_m5) + ggtitle(vsi_v2_m5$formula)

grid.arrange(p1, p2, nrow = 1)

```

```{r, fig.width=12, fig.height=6}

p1 <- stanplot(vsi_v2_m3) + ggtitle(vsi_v2_m4$formula)
p2 <- stanplot(vsi_v2_m4) + ggtitle(vsi_v2_m5$formula)

grid.arrange(p1, p2, nrow = 1)

```

In summary adding, y ~ 1 + (scaled_x | grps) and y ~ 1 + (1 + scaled_x | grps); yield the same set of features, both of them model varying slopes as well as intercepts.

```{r}
pp_check(vsi_v2_m2, type = "intervals_grouped", group = "grps") + ggtitle(vsi_v2_m2$formula)
pp_check(vsi_v2_m3, type = "intervals_grouped", group = "grps") + ggtitle(vsi_v2_m3$formula)
pp_check(vsi_v2_m4, type = "intervals_grouped", group = "grps") + ggtitle(vsi_v2_m4$formula)
pp_check(vsi_v2_m5, type = "intervals_grouped", group = "grps") + ggtitle(vsi_v2_m5$formula)

```

```{r}

mdf %>%
  data_grid(scaled_x, grps) %>%
  add_predicted_draws(vsi_v2_m5) %>%
  group_by(scaled_x, grps) %>%
  sample_n(100) %>%
  ungroup() %>%
  ggplot(aes(x = scaled_x, y = .prediction, col = grps)) +
    geom_point() +
    ggtitle("Predicted Draws")

mdf %>%
  select(scaled_x, grps) %>%
  add_fitted_draws(vsi_v2_m5) %>%
  group_by(scaled_x, grps) %>%
  sample_n(100) %>%
  ungroup() %>%
  ggplot(aes(x = scaled_x, y = .value, col = grps)) +
    geom_point() +
    ggtitle("Fitted Draws")

```

Undertanding the parameters

```{r}

mcmc_intervals(vsi_v2_m5, vars(-lp__, -contains("prior")))
mcmc_intervals_data(vsi_v2_m5, vars(-lp__, -contains("prior")))
```

```{r}

```

